---
title: "`r paste('Azimuth Effect Correction - Week', params$week, 'Hour', params$hour, 'Radar', params$radar)`"
site: "bookdown::bookdown_site"
output:
  bookdown::gitbook:
    code_folding: hide
documentclass: book
params:
  week: 35
  hour: 1
  radar: 'KRIW'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/clo/bigbird/wgfd")

library(bioRad)
library(stringr)
library(lubridate)
library(rhdf5)
library(units)
library(sf)
library(gridExtra)
library(ggplot2)
library(viridisLite)
library(mgcv)
library(tools)
library(dplyr)

```

# Azimuth effect correction

```{r, include=FALSE, eval=TRUE}

this.week <- params$week
this.hour <- params$hour
this.radar <- params$radar


cr = read.csv(paste0('closest_records_', this.radar, '.csv'))
cr <- cr %>%
  mutate(basename = file_path_sans_ext(closest_record))

cr_selected <- cr %>%
  select(basename, hours_after_sunset)

composite_ppi_list <- list.files(
  path = file.path("composite_ppi", "composite", this.radar),
  pattern = "\\.Rds$",
  recursive = FALSE,
  full.names = TRUE
)

composite_file_info <- data.frame(
  filename = basename(composite_ppi_list),
  radar = str_extract(basename(composite_ppi_list), "^[A-Z]+"),
  week = as.numeric(str_extract(str_extract(basename(composite_ppi_list), "_week\\d+"), "\\d+")),
  hour = as.numeric(str_extract(basename(composite_ppi_list), "(?<=_)\\d+(?=_V)")),
  stringsAsFactors = FALSE
)

this.composite <- composite_file_info[composite_file_info$week == this.week & composite_file_info$hour == this.hour,]$filename

kcys_vp_path <- paste0("vp/", this.radar, "/")
kcys_vp_files <- list.files(path = kcys_vp_path, pattern = "\\.h5$", full.names = TRUE, recursive = TRUE)

kcys_file_info <- data.frame(
  full_path = kcys_vp_files,
  filename = basename(kcys_vp_files),
  radar = substr(basename(kcys_vp_files), 1, 4),
  year = as.numeric(substr(basename(kcys_vp_files), 5, 8)),
  month = as.numeric(substr(basename(kcys_vp_files), 9, 10)),
  day = as.numeric(substr(basename(kcys_vp_files), 11, 12)),
  stringsAsFactors = FALSE
)

kcys_file_info$date <- as.Date(with(kcys_file_info, paste(year, month, day, sep = "-")), "%Y-%m-%d")
kcys_file_info$week <- isoweek(as.Date(kcys_file_info$date))
kcys_file_info <- kcys_file_info %>%
  mutate(basename = file_path_sans_ext(filename))

#convert the year, month, day into date object
kcys_file_info$date <- as.Date(with(kcys_file_info, paste(year, month, day, sep = "-")), "%Y-%m-%d")
kcys_file_info$week <- isoweek(as.Date(kcys_file_info$date))

merged_data <- inner_join(kcys_file_info, cr_selected, by = "basename")

matching_vp_files <- list()

matched_files <- merged_data[
      merged_data$week == this.week &
      merged_data$hour == this.hour &
      merged_data$radar == this.radar, 
      "filename"
    ]

matching_vp_files[[this.composite]] <- matched_files

```

## Collect vertical profile statistics
Use `integrate_profile()` to store density `vid` and direction `dd` from each vertical profile of the composite.

```{r, eval=TRUE}

print(matching_vp_files)
print(length(matching_vp_files))

# Get the name of the first list element
first_key <- names(matching_vp_files)[1]

# If there is a name retrieved and it corresponds to a valid entry in the list
if (!is.null(first_key)) {
  # Remove the extension, add '_adjusted', and then re-add the extension
  base_name <- tools::file_path_sans_ext(first_key)
  extension <- tools::file_ext(first_key)
  adjusted_filename <- paste0(base_name, "_adjusted.", extension)

  # Now adjusted_filename holds the new filename with '_adjusted' before the suffix
  print(adjusted_filename)  # Print or use as needed
} else {
  cat("No files in the list.\n")
}


vp_stat_list <- vector("list", length(matching_vp_files))
names(vp_stat_list) <- names(matching_vp_files)

for (i in 1:length(matching_vp_files)) { #just run the first composite for now
  for (j in 1:length(matching_vp_files[[i]])) {
    tryCatch({
      #print(paste("Processing file", j, "in list", i))
      filename <- paste0('vp/', this.radar, '/', basename(matching_vp_files[[i]][j]))
      my_vp <- read_vpfiles(filename)

      if (!is.null(my_vp)) {
        vp_stat_list[[i]]$vid[[j]] <- integrate_profile(my_vp)$vid
        vp_stat_list[[i]]$dd[[j]]  <- integrate_profile(my_vp)$dd
      }
      rhdf5::h5closeAll()
    }, error = function(e) {
      # Error handling
      cat("Error in processing file", j, "in list", i, ": ", e$message, "\n")
    })
  }
}

vp_df_list <- list()

for (composite_name in names(vp_stat_list)) { #just run the first composite for now
  stats <- vp_stat_list[[composite_name]]

  if (length(stats$vid) > 0 && length(stats$dd) > 0) {
    df <- data.frame(
      vid = unlist(stats$vid),  
      dd = unlist(stats$dd),    
      stringsAsFactors = FALSE  
    )
    vp_df_list[[composite_name]] <- df
  } else {
    vp_df_list[[composite_name]] <- data.frame(vid = numeric(0), dd = numeric(0))
  }
}

names(vp_df_list)
head(vp_df_list[[1]])
```

## Compute mean ground speed direction

Adjust `dd` angles by adding 360° to values less than 180° to handle the circular nature of azimuths, then convert degrees to radians. Obtain weighted cosine and sine using `vid`, then use `atan2()` to calculate the mean angle in radians.  After converting back to degrees, use the mean angle as the `reference_angle` in the following step.


```{r, eval=TRUE}

for (name in names(vp_df_list)) {
  df <- vp_df_list[[name]]
  df <- na.omit(df) #remove rows with NA values

  #adjust for circular statistics to avoid problems at 0/360
  df$dd_adjusted <- ifelse(df$dd < 180, df$dd + 360, df$dd) 
  df$radians <- df$dd_adjusted * pi / 180

  weighted_cos <- sum(cos(df$radians) * df$vid) / sum(df$vid)
  weighted_sin <- sum(sin(df$radians) * df$vid) / sum(df$vid)

  mean_angle_radians <- atan2(weighted_sin, weighted_cos)
  mean_angle_degrees <- mean_angle_radians * 180 / pi
  mean_angle_degrees <- ifelse(mean_angle_degrees < 0, mean_angle_degrees + 360, mean_angle_degrees)

  results <- list(mean_angle_degrees = mean_angle_degrees)
  vp_df_list[[name]] <- list(data = df, stats = results)
}

vp_df_list[[1]]$stats$mean_angle_degrees
```

## Compute relative azimuth

The mean direction of bird movement - the reference angle - now serves as 0°. To calculate relative azmiuth for each grid cell, adjust its original azmiuth by the reference angle.  Add `relative_azimuth` to the `ppi` object. 

```{r, message=FALSE, warning=FALSE, results='hide'}

#radar location
lon=my_vp$attributes$where$lon
lat=my_vp$attributes$where$lat

add_azimuth_degrees <- function(spatial_grid, lon, lat, reference_angle = 0) {
  if (!inherits(spatial_grid, "SpatialGridDataFrame")) {
    stop("Input must be a SpatialGridDataFrame.")
  }

  #convert input degrees to radians
  lon_center = lon * pi / 180
  lat_center = lat * pi / 180

  coords <- coordinates(spatial_grid)
  lon_coords = coords[, 1] * pi / 180
  lat_coords = coords[, 2] * pi / 180

  azimuth_radians <- atan2(sin(lon_coords - lon_center),
                           cos(lat_center) * tan(lat_coords) - sin(lat_center) * cos(lon_coords - lon_center))

  azimuth_degrees <- azimuth_radians * 180 / pi
  spatial_grid@data$azimuth <- ifelse(azimuth_degrees < 0, azimuth_degrees + 360, azimuth_degrees)

  azimuth_degrees <- ifelse(azimuth_degrees < 0, azimuth_degrees + 360, azimuth_degrees)
  azimuth_degrees <- (azimuth_degrees - reference_angle + 360) %% 360
  spatial_grid@data$relative_azimuth <- azimuth_degrees

  return(spatial_grid)
}

mean_composite_ppi <- readRDS(paste0('composite_ppi/composite/', this.radar, '/', this.composite)) 

basemap <- rosm::osm.types()[1]

mean_composite_ppi$data <- add_azimuth_degrees(mean_composite_ppi$data
                                               ,lon=lon
                                               ,lat=lat
                                               ,reference_angle=vp_df_list[[1]]$stats$mean_angle_degrees) #change


png_filename <- paste0("week", this.week, "_hour", this.hour, "_composite_unadjusted.png")
png(filename = png_filename, width = 3840, height = 2160, res = 300)
plot_title <- paste0("week ", this.week, " | hour ", this.hour, " | composite")
map(mean_composite_ppi, map = basemap, param = 'VID', zlim = c(0, 100)) + ggtitle(plot_title)
dev.off()
map(mean_composite_ppi, map = basemap, param = 'VID', zlim = c(0, 100)) + ggtitle(plot_title)


# Function to extract the legend from a ggplot
get_legend <- function(my_plot) {
  tmp <- ggplot_gtable(ggplot_build(my_plot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}

basemap <- rosm::osm.types()[1]

# Create the two plots with titles
p1 <- map(mean_composite_ppi, map = basemap,
          alpha = 0.7,
          param='azimuth', zlim=c(0,360), palette = rev(mako(256))) +
  ggtitle("azimuth") +
  theme(legend.position = "none")

p2 <- map(mean_composite_ppi, map = basemap, 
          alpha = 0.7,
          param='relative_azimuth', zlim=c(0,360), palette = rev(mako(256))) +
  ggtitle("relative azimuth") +
  theme(legend.position = "none")

legend <- get_legend(
  map(mean_composite_ppi, param='azimuth', zlim=c(0,360), palette = rev(mako(256))) +
    theme(legend.position = "top", 
          legend.margin = margin(t = -10, r = 0, b = 0, l = 0),  # Adjust top margin to reduce space
          legend.box.margin = margin(t = -10, r = 0, b = 0, l = 0))  # Adjust box margin to reduce space
)

grid.arrange(
  arrangeGrob(p1, p2, ncol = 2),
  legend,
  ncol = 1,
  heights = c(10, 1)  # Keep this ratio; it's less relevant for controlling the gap
)

```

## Compute shifted azimuth

Fit GAM to VID over relative_azimuth and find the peaks of the fitted values. The difference between 90° and the first peak
should give you the value by which to shift your relative azmiuth. The purple line in the plot below shows the final azimuth to be used before mirroring.

```{r, message=FALSE, warning=FALSE, results='hide'}

mirror_azimuth <- function(azimuth) {
  ifelse(azimuth > 180, 360 - azimuth, azimuth)
}

shift_azimuth <- function(azimuth, degrees, x_max = 360) {
  shifted_azimuth <- (azimuth + degrees) %% x_max
  ifelse(shifted_azimuth < 0, shifted_azimuth + x_max, shifted_azimuth)
}

### only take values within 50 KM range
distance_threshold <- set_units(50000, "m")
sf_data <- st_as_sf(mean_composite_ppi$data)
st_crs(sf_data) <- 4326
sf_data_utm <- st_transform(sf_data, 26913)

radar_station <- st_sfc(st_point(c(lon, lat)), crs = 4326)
radar_station_utm <- st_transform(radar_station, 26913)
distances <- st_distance(sf_data_utm, radar_station_utm)
within_50km <- sf_data_utm[distances <= distance_threshold,]

data_for_plot <- data_for_plot_mirror <-  as.data.frame(na.omit(within_50km))
data_for_gam <- data_for_gam_mirror <- na.omit(data_for_plot[, c("VID", "relative_azimuth")])

gam_fit <- gam(VID ~ s(relative_azimuth, bs = "cc", k=6), data = data_for_gam)
summary(gam_fit)
data_for_plot$fitted_values <- predict(gam_fit, newdata = data_for_plot, type = "response")


#plot original VID
ggplot(data_for_gam, aes(x = relative_azimuth, y = VID)) +
  geom_point(alpha=0.1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "VID ~ azimuth (50 km range)",
       x = "Azimuth", y = "VID") +
  theme_minimal()
```



```{r, message=FALSE, warning=FALSE}

# Filter data into two segments
before_180 <- na.omit(data_for_plot[data_for_plot$relative_azimuth < 180, ])
after_180 <- na.omit(data_for_plot[data_for_plot$relative_azimuth >= 180, ])

# Find the peak VID value before azimuth 180
if (nrow(before_180) > 0) {
  first_peak_value <- max(before_180$fitted_values)
  first_peak_index <- which.max(before_180$fitted_values)
  first_peak_azimuth <- before_180$relative_azimuth[first_peak_index]
} else {
  first_peak_value <- NA
  first_peak_azimuth <- NA
}

# Find the peak VID value after azimuth 180
if (nrow(after_180) > 0) {
  second_peak_value <- max(after_180$fitted_values)
  second_peak_index <- which.max(after_180$fitted_values)
  second_peak_azimuth <- after_180$relative_azimuth[second_peak_index]
} else {
  second_peak_value <- NA
  second_peak_azimuth <- NA
}

# Calculate the shift degrees for both peaks
shift_degrees_first_peak <-  90 - first_peak_azimuth
shift_degrees_second_peak <- 270 - second_peak_azimuth

# Compute the average shift degrees
shift_degrees <- mean(c(shift_degrees_first_peak, shift_degrees_second_peak))

# Output the peaks and calculated shift degrees
cat("first peak: VID =", first_peak_value, ", azimuth =", first_peak_azimuth, "\n")
cat("second peak: VID =", second_peak_value, ", azimuth =", second_peak_azimuth, "\n")
cat("average shift degrees:", shift_degrees, "\n")


data_for_plot$shifted_azimuth <- sapply(data_for_plot$relative_azimuth, shift_azimuth, degrees = shift_degrees)


#mirror the shifted values
data_for_gam_mirror$mirrored_azimuth <- sapply(data_for_plot$shifted_azimuth, mirror_azimuth)
gam_fit <- gam(VID ~ s(mirrored_azimuth, bs = "cc", k=6), data = data_for_gam_mirror)
summary(gam_fit)
data_for_gam_mirror$fitted_values <- predict(gam_fit, newdata = data_for_gam_mirror, type = "response")


peak_difference <- round(second_peak_azimuth - first_peak_azimuth, 2)
center_azimuth <- (first_peak_azimuth + second_peak_azimuth) / 2
center_value <- (first_peak_value + second_peak_value) / 2


plot <- ggplot(data_for_plot, aes(x = relative_azimuth, y = VID)) +
  geom_point(aes(color = "Observed"), alpha = 0.05, size = 0.5) +
  geom_line(aes(y = fitted_values, color = "Fitted"), size = 1) +
  geom_line(aes(x = shifted_azimuth, y = fitted_values, color = "Shifted"), linetype = "dashed", size = 1) +
  labs(x = "", y = "VID", title = "GAM fit VID ~ Azimuth") +
  scale_color_manual(name = "", values = c("Observed" = "blue", "Fitted" = "red", "Shifted" = "purple")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_text(size = 0, margin = margin(t = 0, b = 0, l = 0, r = 0))) +  # Minimize x-axis label space
  geom_text(aes(x = first_peak_azimuth, y = first_peak_value, label = paste(round(first_peak_azimuth), "°", sep=" ")),
            vjust = -1, color = "black") +
  geom_text(aes(x = second_peak_azimuth, y = second_peak_value, label = paste(round(second_peak_azimuth), "°", sep=" ")),
            vjust = -1, color = "black") +
  geom_segment(aes(x = first_peak_azimuth, y = first_peak_value, xend = second_peak_azimuth, yend = second_peak_value),
               linetype = "dashed", color = "black") +
  ylim(0, 100) +
  geom_vline(xintercept = 90, linetype = "solid", color = "black", size = 1, alpha = 0.4) +
  geom_vline(xintercept = 270, linetype = "solid", color = "black", size = 1, alpha = 0.4) +
  geom_text(aes(x = center_azimuth, y = center_value, label = paste(peak_difference, "°")),
            vjust = 2, color = "black")
  ylim(0, 100)
  
plot
```

## Mirror azimuth

Mirror azmiuths across the 180° axis

```{r, message=FALSE, warning=FALSE}

#plot mirrored VID
ggplot(data_for_gam_mirror, aes(x = mirrored_azimuth, y = VID)) +
  geom_point(alpha=0.1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "VID ~ azimuth (50 km range)",
       x = "Azimuth", y = "VID") +
  theme_minimal()

```

## Normalize VID

First center the fitted VID values around zero by subtracting the mean, then normalize between -1 and 1 by dividing by the maximum absolute value. The net area under the normalized curve is calculated to confirms correct normalization (should be close to 0). 

```{r}

# center the data around zero
centered_VID <- data_for_gam_mirror$fitted_values - mean(data_for_gam_mirror$fitted_values)

# normalize the data to range -1 to 1
max_abs_value <- max(abs(centered_VID))
normalized_VID <- centered_VID / max_abs_value

# add normalized data to thae dataframe
data_for_gam_mirror$normalized_VID <- normalized_VID

# calculate the net area to verify it's close to zero
azimuth_diff <- mean(diff(data_for_gam_mirror$mirrored_azimuth))
net_area <- sum(normalized_VID * azimuth_diff)

# Plot the normalized VID values
ggplot(data_for_gam_mirror, aes(x = mirrored_azimuth, y = normalized_VID)) +
  geom_line() +  # Draw the line
  geom_area(aes(y = ifelse(normalized_VID > 0, normalized_VID, 0)), fill = "blue", alpha = 0.1) +  # Fill area above zero
  geom_area(aes(y = ifelse(normalized_VID < 0, normalized_VID, 0)), fill = "red", alpha = 0.1) +  # Fill area below zero
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  # Add a horizontal line at y = 0
  labs(title = "Normalized VID across azimuth",
       x = "Azimuth (degrees)", y = "Normalized VID") +
  theme_minimal()

# Output the net area
print(paste("Net area under the curve: ", net_area))

```

## Compute VID adjustment factor

The total sum of the adjusted VID values must remain consistent with the original sum. We calculate the proportion of positive and negative adjustments based on the sum of VID values for data points with normalized VID values greater than and less than or equal to zero. Then we apply adjustment to each normalized VID value, reducing positive values and boosting negative ones. Finally, a scaling factor is computed by dividing the original sum of VID values by the sum of the adjusted VID values, ensuring the total remains unchanged.

```{r, message=FALSE, warning=FALSE}

valid_data <- data_for_gam_mirror[!is.nan(data_for_gam_mirror$VID), ]
total_VID <- sum(valid_data$VID)

positive_adjustment_sum <- sum(valid_data$VID[valid_data$normalized_VID > 0])
negative_adjustment_sum <- sum(valid_data$VID[valid_data$normalized_VID <= 0])

prop_positive <- ifelse(positive_adjustment_sum > 0, positive_adjustment_sum / total_VID, 0)
prop_negative <- ifelse(negative_adjustment_sum > 0, negative_adjustment_sum / total_VID, 0)

valid_data$adjustment_factor <- ifelse(valid_data$normalized_VID > 0,
                                       1 - prop_positive * abs(valid_data$normalized_VID),
                                       1 + prop_negative * abs(valid_data$normalized_VID))

valid_data$adjusted_VID <- valid_data$VID * valid_data$adjustment_factor

adjusted_sum <- sum(valid_data$adjusted_VID)
original_sum <- sum(valid_data$VID)

scaling_factor <- original_sum / adjusted_sum
valid_data$adjustment_factor <- valid_data$adjustment_factor * scaling_factor
valid_data$adjusted_VID <- valid_data$VID * valid_data$adjustment_factor
data_for_plot_mirror$adjusted_VID[!is.nan(data_for_plot_mirror$VID)] <- valid_data$adjusted_VID

sum_adjusted_VID <- sum(valid_data$adjusted_VID)

ggplot(valid_data, aes(x = mirrored_azimuth, y = adjustment_factor)) +
  geom_line(color = "green", size = 1) +
  labs(title = "Adjustment factor ~ azimuth",
       x = "azimuth",
       y = "adjustment factor") +
  theme_minimal()


range_data <- apply(valid_data, 2, range)
print(range_data)

```


```{r, message=FALSE, warning=FALSE}

azimuth <- valid_data$mirrored_azimuth
adjustment_factor <- valid_data$adjustment_factor
new_azimuth <- seq(0,180, length.out = 100)
interpolated_values <- approx(x = azimuth, y = adjustment_factor, xout = new_azimuth)
new_adjustment_factor <- interpolated_values$y

plot(azimuth, adjustment_factor, main="Approximation of Adjustment Factor vs. Mirrored Azimuth",
     xlab="Mirrored Azimuth", ylab="Adjustment Factor", pch=16, col="blue")
lines(new_azimuth, new_adjustment_factor, col="red", lwd=2)
legend("topright", legend=c("Original Data", "Interpolated Line"), col=c("blue", "red"), pch=c(16, NA), lty=c(NA, 1), lwd=c(NA, 2))


#shift and mirror azimuth values in composite ppi
mean_composite_ppi$data$shifted_azimuth = sapply(mean_composite_ppi$data$relative_azimuth, shift_azimuth, degrees = shift_degrees)
mean_composite_ppi$data$mirrored_azimuth = sapply(mean_composite_ppi$data$shifted_azimuth, mirror_azimuth)

mean_composite_ppi$data$adjustment_factor <- approx(
  x = valid_data$mirrored_azimuth,
  y = valid_data$adjustment_factor,
  xout = mean_composite_ppi$data$mirrored_azimuth,
  rule = 2
)$y

```



```{r, message=FALSE, warning=FALSE}

output_directory <- paste0("composite_ppi/composite/", this.radar, "/corrected")
images_directory <- paste0("composite_ppi/composite/", this.radar, "/corrected/images")

output_path <- file.path(output_directory, paste0(adjusted_filename, ".Rds"))
plot_title <- sprintf("week %s | hour %s | composite", this.week, this.hour)
png_filename <- sub("\\.Rds$", ".png", adjusted_filename)
png_full_path <- file.path(images_directory, png_filename)


mean_composite_ppi$data$adjusted_VID <- ifelse(
  is.na(mean_composite_ppi$data$VID), 
  NA, 
  mean_composite_ppi$data$VID * mean_composite_ppi$data$adjustment_factor
)

map(mean_composite_ppi, map=basemap, param='adjusted_VID', zlim=c(0,100)) + ggtitle(plot_title)
saveRDS(mean_composite_ppi, file = output_path)

png(filename = png_full_path, width = 3840, height = 2160, res = 300)
map(mean_composite_ppi, map=basemap, param='adjusted_VID', zlim=c(0,100)) + ggtitle(plot_title)
dev.off()

```







